On the Uncertainty of Wind Power Generation

### 2019-10-01 LaTeX_code/presentation/DELFT_pres.tex
lamperti optimization adjustment

\title{On the Uncertainty of Wind Power Generation}
% \subtitle{ Waleed Alhaddad \ Raul Tempone \ Ahmad kebaier }

\author{ Waleed Alhaddad\textsuperscript{\textasteriskcentered} \qquad Ahmed Kebaier\textsuperscript{\ddag} \qquad Ra\'ul  Tempone\textsuperscript{\textasteriskcentered}\textsuperscript{\textdagger} \\  \vskip 0.2in
\textsuperscript{\textasteriskcentered}CEMSE Division, KAUST , Saudi Arabia
 \\ \textsuperscript{\ddag}Université Paris 13, Sorbonne Paris Cité, LAGA, CNRS (UMR 7539) , France  \\ \textsuperscript{\textdagger}Alexander von Humboldt Professor, RWTH Aachen University,  Germany}

\begin{document}
\setbeamercolor{background canvas}{bg=blue!10}

\begin{frame}
\titlepage
\end{frame}

@@

\begin{frame}\frametitle{Status quo }
  Wind power forecasts can be generally categorized as follows:
  \begin{itemize}
    \item physical models
    \item statistical methods
    \item artificial intelligence methods
    \item spatial correlation methods
    \item persistence models
    \item other hybrid approaches
  \end{itemize}
 The output of such methods is usually a \textbf{deterministic forecast}. Occasionally probabilistic forecasts are produced through uncertainty propagation in the data, parameters or through forecast ensembles. However, little has been done in terms of producing \textbf{data driven probabilistic forecasts} based on the real-world performance of forecasting models.
\end{frame}

\begin{frame}\frametitle{Data}
This is a year long data set from Uruguay based on \textbf{10 minute observation interval of one thousand 72-hour paths} (2018).
\begin{figure}
  \includegraphics[width=70mm,scale=1]{plots/data_1516064400.pdf}
  \includegraphics[width=70mm,scale=1]{plots/data_1516323600.pdf}\\
  \includegraphics[width=70mm,scale=1]{plots/data_1517619600.pdf}
  \includegraphics[width=70mm,scale=1]{plots/data_1529456400.pdf}
  % \caption{title}
\end{figure}
% Do you have more data? send it our way.
\end{frame}

\begin{frame}\frametitle{Data Skewness}
\begin{figure}
  \includegraphics[width=47.5mm,scale=1]{plots/hist_low.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_mid.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_high.pdf}
  \caption{We see that forecast errors exhibit skewness near the boundries (i.e. low and high power production regimes.)}
\end{figure}
\end{frame}

@@

\begin{frame}\frametitle{Model}
We propose to model wind power\textbf{ forecasts errors using parametric stochastic differential equations (SDEs)} whose solution defines a stochastic process. This resultant stochastic process describes the time evolution dynamics of wind power forecast errors.
\begin{equation}
\begin{split}
dX_t &= a(X_t; \bm{\theta}) dt + b (X_t; \bm{\theta} ) dW_t \quad t > 0 \\
X_0 & = X_0
\end{split}
\label{main}
\end{equation}

\begin{itemize}
\item $a(\cdot; \bm{\theta}):[0,1] \to \R $  a drift function.
\item $b (\cdot; \bm{\theta} ):[0,1] \to \R$  a  diffusion function.
\item $\bm{\theta}$: a vector of parameters.
\item $W_t$: Standard Wiener random process in $\R$.
\end{itemize}

Question: How do we choose an appropriate drift and diffusion functions?

@@

\begin{frame}\frametitle{Model}
Answer:
\begin{enumerate}
  \item We want the process to follow the wind forecast, thus we choose a drift term that is mean reverting and tracks the derivative of the deterministic forecast $p$, which is an input to our model.
  \begin{equation}
    a(X_t; \bm{\theta})=  \dot{p} \ dt - \theta_t(X_t - p_t)
  \end{equation}
where $\theta_t$ is a time-dependent paramter that controls the speed of reversion.
  \item We want a diffusion term that vanishes at the boundaries to prevent the process from escaping the region $[0,1]$.
  \begin{equation}
    b (\cdot; \bm{\theta} )= \sqrt{2 \theta_t \alpha x (1-x)}
  \end{equation}
  where $\alpha$ is a constant prameter that controls the path variability.

To further ensure that the process does not escape the region $[0,1]$, the mean reversion parameter has to be selected according to the following rule,
\begin{equation}
\theta_t = \max \left( \theta_0 \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right ) \label{theta_t}
\end{equation}
\end{enumerate}
\end{frame}

\begin{frame}\frametitle{Model}
Thus, our SDE becomes
\begin{equation}
\begin{split}
dX_t&= \dot{p}_t \ dt - \theta_t(X_t - p_t) \ dt + \sqrt{2 \theta_t \alpha x (1-x)}  \ dW_t \quad t > 0 \\
X_0&=x_0\\
% \theta_t &= \max \left( \theta_0 \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right )\\
\end{split}
\label{model:derivative_tracking_X}
\end{equation}
To avoid differentiation of the forecast $p_t$ and simplify, we apply a change of variables $$V_t = X_t - p_t$$ \\
The  model becomes,
\begin{equation}
\begin{split}
dV_t &=  - \theta_t V_t \  dt + \sqrt{2 \theta_t \alpha (V_t +p_t ) (1-V_t-p_t)} \  dW_t  \\ %\quad t > 0
V_0 & = v_0\\
% \theta_t &= \max \left( \theta_0 \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right )\\
\end{split}\label{VtSDE}
\end{equation}
Note that this model is Markovian.
\end{frame}

\begin{frame}\frametitle{Model}
Since $V_t$ defined by our SDE is Markovian, the likelihood function can be written as a product of transition densities.

@@

\begin{equation}
\begin{split}
\frac{ \partial f }{\partial t } & (y ,t | x , s, \theta_t, \alpha )= - \frac{\partial}{ \partial y} ( a( y;\dot{p}_t , p_t, \theta_t ) f( y ,t | x , s, \theta_t , \alpha ) ) \\
& + \frac{1}{2} \frac{\partial^2}{ \partial y^2} ( b(y;\theta_t, \alpha  )  f( y ,t | x , s, \theta_t , \alpha )  ) \quad  t < s\\
\end{split}
\end{equation}
This is a parametric PDE which is computationally prohibitive to solve and optimize for every transition.
\end{frame}

\begin{frame}\frametitle{Moment Matching}
Instead of solving for exact transition densities by the Fokker-Planck, we propose a \textbf{proxy transition denisty}. We match the moments of our SDE model with that of the proxy density. Using Ito, we arrive at the following iterative ODEs.

\begin{equation}
\begin{split}
\frac{d \E[ V^k_t]}{dt} = - k \theta_t \E [ V^k_t] + \frac{k(k-1)}{2} \E [ V^{k-2}_t  b(y;\theta_t, \alpha)]
\end{split}
\end{equation}
For the first two moments, we obtain
\begin{equation}
\begin{split}
\frac{d m_1 (t)}{dt} &= - m_1(t)\theta_t \\
\frac{d m_2 (t)}{dt} &=  -2 m_2(t)\theta(1+\alpha) + 2\alpha\theta m_1(t)(1-2p_t) + 2 \alpha\theta p_t (1-p_t)\\
\end{split}
\end{equation}

 A suitable candidate is a Beta transition density as it is compactly supported and can morph into symmetric and asymmetric shapes.

@@

Execute the following until an accuracy threshold is met:
\begin{enumerate}
\item initialize.
\item optimize the log-likelihood function.
\begin{enumerate}
\item For every evaluation of the log-liklihood function:
\begin{enumerate}
  \item Sample a batch of transitions and their associated forecast and parameters.
  \item Solve the ODE system for every transition.
  \item Match the resulting moments with the parameters of the chosen proxy distribution for every transition.
\end{enumerate}
\end{enumerate}
\item go to step 1 (i.e. re-initialize the optimization with the most recent result ).
\end{enumerate}

Notes:
\begin{itemize}
\item Choose your favorite deterministic optimization algorithm.
\item Choose your favorite integrator to solve the ODE system.
\end{itemize}
\end{frame}

\begin{frame}\frametitle{Inference Results}
Here we show the convergence path of the optimization in 2D, the likelihood in 3D and rate of the shrinking ellipses in one slide.
\end{frame}

\begin{frame}\frametitle{Simulation and Confidence Bands}
Here we show simulated future wind power production and another plot with confidence bands.
\end{frame}

\begin{frame}\frametitle{State-Independent model}
The model we have demostrated previously is state-dependent, that is the diffusion of the SDE depends on the state.
\begin{equation}
\begin{split}
dV_t &=  - \theta_t V_t \  dt + \sqrt{2 \theta_t \alpha (V_t +p_t ) (1-V_t-p_t)} \  dW_t  \\ %\quad t > 0
V_0 & = v_0\\
\end{split}\label{VtSDE}
\end{equation}

@@

\begin{equation}
  \begin{split}
    dZ_t&= \frac{- \theta_t (1+ \sin(Z_t) - 2p_t) + \alpha \theta_t \sin (Z_t)   }{\cos (Z_t)} + \sqrt{2 \alpha \theta_t} dW_t \\
    Z_0&=z_0
  \end{split}
\end{equation}
where $Z_t = \arcsin \left( \frac{1}{2} \left( V_t+p_t \right) - 1 \right) $.
\end{frame}

\begin{frame}\frametitle{Data Skewness after Lamperti transformation}
As stated before, the state-independent SDE follows a Lamperti transformed process $Z_t$ give by $Z_t = \arcsin \left( \frac{1}{2} \left( V_t+p_t \right) - 1 \right) $.
\begin{figure}
  \includegraphics[width=50mm,scale=1]{plots/hist_low_lamperti.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_mid_lamperti.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_high_lamperti.pdf}
  \caption{We observe that most of the skewness has been removed after the lamperti transformation. This motivates us to use a Gaussian transition density as a proxy denisty.}
\end{figure}
\end{frame}

\begin{frame}\frametitle{State-Independent model}
Similarly, we try to obtain a system of ODEs to determine the centered moments of the Lamperti transformed process $V_t$. Due to the non-linearity in the drift, we can only approximate the centered moments by the following ODEs,
\begin{equation}
\begin{split}
\frac{d m_1 (t)}{dt} &= - m_1(t)\theta_t (1-\alpha) - \theta (1-2 p_t) \\
\frac{d v(t)}{dt} &=  2 v(t) \theta_t (2p_t - 1 ) \tan(x) \sec(x) + \theta_t (\alpha - 1) \sec^2(x)  + 2 \theta_t \alpha\\
\end{split}
\end{equation}
These are not exact moment ODEs, however they are good enough for small time intervals.
\end{frame}

\begin{frame}\frametitle{Inference Results}
Here we show the convergence path of the optimization in 2D, the likelihood in 3D and rate of the shrinking ellipses in one slide.
\end{frame}

\begin{frame}\frametitle{Simulation and Confidence Bands}
Here we show simulated future wind power production and another plot with confidence bands.
\end{frame}

\begin{frame}\frametitle{Model comparison}
\begin{table}[]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
Model   &  parameters $(\theta_0, \alpha)$   & AIC & BIC \\ \hline
state-dependent &   (,)  &     &      \\ \hline
state-independent &   (,)  &     &       \\ \hline
\end{tabular}
\caption{We compare the two models based on information criterion here.... }
\label{tab:model_comparison}
\end{table}
\end{frame}

\begin{frame}\frametitle{Concluding remarks}
We were able to:
\begin{itemize}
  \item simulate future wind power production based on real data.
  \item have an analytical description of the uncertainty of wind power forecasts in the form of an SDE.
  \item develop a forecasting technology agnostic method.
  \item capture skewness of the error process and its correlation structure.
\end{itemize}
\end{frame}

\begin{frame}\frametitle{References}
\begin{itemize}
  \item source
  \item source
  \item source
\end{itemize}
\end{frame}

>>>>

\title{On the Uncertainty of Wind Power Generation}

\author{ Waleed Alhaddad\textsuperscript{\textasteriskcentered} \qquad Ahmed Kebaier\textsuperscript{\ddag} \qquad Ra\'ul  Tempone\textsuperscript{\textasteriskcentered}\textsuperscript{\textdagger} \\  \vskip 0.2in
\textsuperscript{\textasteriskcentered}CEMSE Division, KAUST, Saudi Arabia \vskip 0.05in  \textsuperscript{\ddag}Université Paris 13, Sorbonne Paris Cité, LAGA, CNRS (UMR 7539) , France\vskip 0.05in  
 \textsuperscript{\textdagger}Alexander von Humboldt Professor, RWTH Aachen  University,  Germany}

\begin{document}
\setbeamercolor{background canvas}{bg=blue!10}
\date{}

\begin{frame}
\titlepage
\center
  \includegraphics[width=40mm,scale=1]{KAUST_Logo.pdf}\quad 
  \includegraphics[width=60mm,scale=1]{rwth_UQ.png}\quad 
  \includegraphics[width=40mm,scale=1]{UParis_13.png} \quad
\end{frame}

@@

\begin{frame}\frametitle{Status quo }
  Wind power forecasts can be generally categorized as follows:
  \begin{itemize}
    \item physical models
    \item statistical methods
    \item artificial intelligence methods
    \item spatial correlation methods
    \item other hybrid approaches
  \end{itemize}
 The output of such methods is usually a \textcolor{red}{deterministic forecast}. Occasionally probabilistic forecasts are produced through uncertainty propagation in the data, parameters or through forecast ensembles. However, little has been done in terms of producing \textcolor{red}{data driven probabilistic forecasts} based on the real-world performance of forecasting models.
\end{frame}

\begin{frame}\frametitle{Data}
This is a year long data set from Uruguay based on \textbf{1000 72-hour long paths of observations every 10 min \textcolor{red}{ ($\sim$ half a million data points)}} recorded in 2018.
\begin{figure}
  \includegraphics[width=70mm,scale=1]{plots/data_1516064400.pdf}
  \includegraphics[width=70mm,scale=1]{plots/data_1516323600.pdf}\\
  \includegraphics[width=70mm,scale=1]{plots/data_1517619600.pdf}
  \includegraphics[width=70mm,scale=1]{plots/data_1529456400.pdf}
  % \caption{title}
\end{figure}
% Do you have more data? send it our way.
\end{frame}

\begin{frame}\frametitle{Data Skewness}
\begin{figure}
  \includegraphics[width=47.5mm,scale=1]{plots/hist_low.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_mid.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_high.pdf}
  \caption{We see that forecast errors exhibit \textcolor{red}{skewness} near the boundries (i.e. low and high power production regimes.)}
\end{figure}
\end{frame}

@@

\begin{frame}\frametitle{Model}
We propose to model wind power forecasts errors using \textcolor{red}{parametric stochastic differential equations (SDEs)} whose solution defines a stochastic process. This resultant stochastic process describes the time evolution dynamics of wind power forecast errors.
\begin{equation}
\begin{split}
dX_t &= a(X_t; p_t,\bm{\theta}) dt + b (X_t;p_t, \bm{\theta} ) dW_t \quad t > 0 \\
X_0 & = x_0
\end{split}
\label{main}
\end{equation}

\begin{itemize}
\item $a(\cdot; \bm{\theta}):[0,1] \to \R $  a drift function.
\item $b (\cdot; \bm{\theta} ):[0,1] \to \R$  a  diffusion function.
\item $\bm{\theta}$: a vector of parameters.
\item $p_t$ time-dependent scalar value.
\item $W_t$: Standard Wiener random process in $\R$.
\end{itemize}

\textcolor{blue}{Question: How do we choose an appropriate drift and diffusion functions?}

@@

\begin{frame}\frametitle{Model}
\textcolor{blue}{Answer:}
Let $\bm{\theta} = (\textcolor{red}{\theta_0}, \textcolor{red}{\alpha})$
\begin{enumerate}
  \item We want the process to follow the wind forecast, thus we choose a drift term that is mean reverting and tracks the derivative of the deterministic forecast $p_t$, which is an input to our model.
  \begin{equation}
    a(x; \bm{\theta})=  \dot{p_t} - \theta_t(x - p_t)
  \end{equation}
where $\theta_t$ is a time-dependent parameter that controls the speed of reversion.
  \item We want a diffusion term that vanishes at the boundaries to prevent the process from escaping the region $[0,1]$.
  \begin{equation}
    b (x; \bm{\theta} )= \sqrt{2 \theta_t \textcolor{red}{\textcolor{red}{\alpha}} x (1-x)}
  \end{equation}
  where $\textcolor{red}{\alpha}$ is a constant parameter that controls the path variability.

To further ensure that the process does not escape the region $[0,1]$, the mean reversion parameter has to be selected according to the following rule,
\begin{equation}
\theta_t = \max \left( \textcolor{red}{\theta_0} \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right ) \label{theta_t}
\end{equation}
\end{enumerate}
\end{frame}

\begin{frame}\frametitle{Model}
Thus, our SDE becomes
\begin{equation}
\begin{split}
dX_t&= \dot{p}_t \ dt - \theta_t(X_t - p_t) \ dt + \sqrt{2 \theta_t \textcolor{red}{\alpha} X_t (1-X_t)}  \ dW_t \quad t > 0 \\
X_0&=x_0\\
% \theta_t &= \max \left( \textcolor{red}{\theta_0} \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right )\\
\end{split}
\label{model:derivative_tracking_X}
\end{equation}
To avoid differentiation of the forecast $p_t$ and simplify, we apply a change of variables $$V_t = X_t - p_t$$ \\
The  model becomes,
\begin{equation}
\begin{split}
dV_t &=  - \theta_t V_t \  dt + \sqrt{2 \theta_t \textcolor{red}{\alpha} (V_t +p_t ) (1-V_t-p_t)} \  dW_t  \\ %\quad t > 0
V_0 & = v_0\\
% \theta_t &= \max \left( \textcolor{red}{\theta_0} \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right )\\
\end{split}\label{VtSDE}
\end{equation}
Note that this model is \textcolor{red}{Markovian.}
\end{frame}

\begin{frame}\frametitle{Model}
Since $V_t$ defined by our SDE is Markovian, the \textcolor{red}{ likelihood function can be written as a product of transition densities}.  Consider a set of M paths with N observations each, $ V^{M,N}=\{ V_{t_1^{M,N}} , V_{t_2^{M,N}} ,\ldots , V_{t_N^{M,N}} \}$ observed in intervals of $\Delta_N$.

@@

\begin{equation}
\begin{split}
\frac{ \partial f }{\partial t } & (y ,t | x , s, \theta_t, \textcolor{red}{\alpha} )= - \frac{\partial}{ \partial y} ( a( y;\dot{p}_t , p_t, \theta_t ) f( y ,t | x , s, \theta_t , \textcolor{red}{\alpha} ) ) \\
& + \frac{1}{2} \frac{\partial^2}{ \partial y^2} ( b(y;\theta_t, \textcolor{red}{\alpha}  )  f( y ,t | x , s, \theta_t , \textcolor{red}{\alpha} )  ) \quad  t < s\\
\end{split}
\end{equation}
This is a parametric PDE which is \textcolor{red}{computationally expensive to solve} and optimize for every transition.
\end{frame}

\begin{frame}\frametitle{Moment Matching}
Instead of solving for exact transition densities by the Fokker-Planck, we propose a \textcolor{red}{proxy transition density}. We match the moments of our SDE model with that of the proxy density. Using Ito, we arrive at the following iterative ODEs.

\begin{equation}
\begin{split}
\frac{d \E[ V^k_t]}{dt} = - k \theta_t \E [ V^k_t] + \frac{k(k-1)}{2} \E [ V^{k-2}_t  b(V^k_t;\theta_t, \textcolor{red}{\alpha})]
\end{split}
\end{equation}
For $t\in [t_{n-1}, t]$, the first two moments are given by
\begin{equation}
\begin{split}
\frac{d m_1 (t)}{dt} &= - m_1(t)\theta_t \\
\frac{d m_2 (t)}{dt} &=  -2 m_2(t)\theta(1+\textcolor{red}{\alpha}) + 2\textcolor{red}{\alpha}\theta_t m_1(t)(1-2p_t) + 2 \textcolor{red}{\alpha}\theta_t p_t (1-p_t)\\
\end{split}
\end{equation}
with initial conditions, $m_1(t_{n-1})= v_{n-1}$ and $m_2(t_{n-1})= v_{n-1}^2$ where $t_{n-1}$ is the time of the previous obseravtion and $v_{n-1}$ is its value.

 A suitable candidate is a \textcolor{red}{Beta transition } density as it is compactly supported and can morph into symmetric and asymmetric shapes.

@@

Execute the following until an accuracy threshold is met:
\begin{enumerate}
\item \textcolor{red}{initialize}.
\item \textcolor{red}{optimize} the log-likelihood function.
\begin{enumerate}
\item For every evaluation of the log-liklihood function:
\begin{enumerate}
  \item \textcolor{red}{Sample} a random mini-batch of transitions and their associated forecast and parameters.
  \item \textcolor{red}{Solve} the ODE system for every transition to obtain the moments.
  \item \textcolor{red}{Match} the resulting moments with the parameters of the chosen proxy distribution for every transition.
\end{enumerate}
\end{enumerate}
\item go to step 1 (i.e. \textcolor{red}{re-initialize} the optimization with the most recent result ).
\end{enumerate}

In the above:
\begin{itemize}
\item Choose your favorite deterministic optimization algorithm.
\item Choose your favorite integrator to solve the ODE system.
\end{itemize}
\end{frame}

\begin{frame}\frametitle{Inference Results}
% Here we show the convergence path of the optimization in 2D, the likelihood in 3D and rate of the shrinking ellipses in one slide.

\begin{figure}
  \includegraphics[width=110mm,scale=1]{plots/beta_likeli_3d.pdf}
  \caption{3-D view of the \textcolor{red}{inverted beta log-likelihood}  function of 100 sample paths. That is a total of  \textcolor{red}{ $\sim 42,700$ data points }}
\end{figure}
\end{frame}

\begin{frame}\frametitle{Inference Results}
\begin{figure}
  \includegraphics[width=70mm,scale=1]{plots/alpha_conv_beta.pdf}
  \includegraphics[width=70mm,scale=1]{plots/theta_conv_beta.pdf}
  \caption{We have \textcolor{red}{self-convergence} of our algorithm at a rate that matches the convergence rate of Monte Carlo.}
\end{figure}
\end{frame}

\begin{frame}\frametitle{Future Wind Power Simulation}
    \begin{figure}
      \includegraphics[width=75mm,scale=1]{simulated/24hr/1099.pdf}
      \includegraphics[width=75mm,scale=1]{simulated/24hr/1178.pdf}
      \caption{We simulate five possible future wind power production paths using the obtain optimal parameters $(\textcolor{red}{\theta_0}, \textcolor{red}{\alpha} )=(12,0.3)$ }
    \end{figure}
\end{frame}

\begin{frame}\frametitle{Confidence Bands}
    \begin{figure}
      \includegraphics[width=75mm,scale=1]{confidence_intervals/24hr/31.pdf}
      \includegraphics[width=75mm,scale=1]{confidence_intervals/24hr/820.pdf}
       \caption{We obtain confidence intervals for future wind power production using the obtain optimal parameters $(\textcolor{red}{\theta_0}, \textcolor{red}{\alpha} )=(12,0.3)$. Actual production plotted in retrospect. }
    \end{figure}
\end{frame}

\begin{frame}\frametitle{State-Independent model}
The model we have demonstrated previously is state-dependent, that is the diffusion of the SDE depends on the state.
\begin{equation}
\begin{split}
dV_t &=  - \theta_t V_t \  dt + \sqrt{2 \theta_t \textcolor{red}{\alpha} (V_t +p_t ) (1-V_t-p_t)} \  dW_t  \\ %\quad t > 0
V_0 & = v_0\\
\end{split}\label{VtSDE}
\end{equation}

@@

\begin{equation}
  \begin{split}
    dZ_t&= \frac{- \theta_t (1+ \sin(Z_t) - 2p_t) + \textcolor{red}{\alpha} \theta_t \sin (Z_t)   }{\cos (Z_t)} + \sqrt{2 \textcolor{red}{\alpha} \theta_t} dW_t \\
    Z_0&=z_0
  \end{split}
\end{equation}
where $Z_t = \arcsin \left( \frac{1}{2} \left( V_t+p_t \right) - 1 \right) $.
\end{frame}

\begin{frame}\frametitle{Data Skewness after Lamperti transformation}
As stated before, the state-independent SDE follows a Lamperti transformed process $Z_t$ give by $Z_t = \arcsin \left( \frac{1}{2} \left( V_t+p_t \right) - 1 \right) $.
\begin{figure}
  \includegraphics[width=50mm,scale=1]{plots/hist_low_lamperti.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_mid_lamperti.pdf}
  \includegraphics[width=50mm,scale=1]{plots/hist_high_lamperti.pdf}
  \caption{We observe that skewness has been greatly reduced after the Lamperti transformation. This motivates us to use a \textcolor{red}{Gaussian transition} density as a proxy density.}
\end{figure}
\end{frame}

\begin{frame}\frametitle{State-Independent model}
Similarly, we try to obtain a system of ODEs to determine the centered moments of the Lamperti transformed process $V_t$. Due to the non-linearity in the drift, we can only approximate the centered moments by the following ODEs,
\begin{equation}
\begin{split}
\frac{d m_1 (t)}{dt} &= - m_1(t)\theta_t (1-\textcolor{red}{\alpha}) - \theta (1-2 p_t) \\
\frac{d v(t)}{dt} &=  2 v(t) \theta_t (2p_t - 1 ) \tan(x) \sec(x) + \theta_t (\textcolor{red}{\alpha} - 1) \sec^2(x)  + 2 \theta_t \textcolor{red}{\alpha}\\
\end{split}
\end{equation}
These are not exact moment ODEs, however they are accurate enough for small time intervals.
\end{frame}

\begin{frame}\frametitle{Inference Results}
\begin{figure}
  \includegraphics[width=70mm,scale=1]{plots/alpha_conv_beta.pdf}
  \includegraphics[width=70mm,scale=1]{plots/theta_conv_beta.pdf}
    \caption{We have self-convergence of our algorithm in the Lamperti space at a rate that matches the convergence rate of Monte Carlo.}
\end{figure}
\end{frame}

\begin{frame}\frametitle{Model comparison}
\begin{table}[]
\centering
\begin{tabular}{|c|c|}
\hline
Model   &  parameters $(\textcolor{red}{\theta_0}, \textcolor{red}{\alpha})$    \\ \hline
state-dependent &   (12,0.3)    \\ \hline
state-independent &   (14,0.29)     \\ \hline
\end{tabular}
\caption{We compare the parameters obtain in both the original and Lamperti space. }
\label{tab:model_comparison}
\end{table}
\end{frame}

\begin{frame}\frametitle{Concluding remarks}
We were able to:
\begin{itemize}
  \item simulate future wind power production based on real data.
  \item obtain an analytical description of the uncertainty of wind power forecasts in the form of an SDE.
  \item develop a forecasting technology agnostic method.
  \item capture skewness of the error process and its correlation structure.
\end{itemize}
\end{frame}

\begin{frame}\frametitle{References}

\section*{References}
\begin{flushleft}
\begin{itemize}
\item Alhaddad, W. ,  Kebaier, A. , \& Tempone, R. (2019). Stochastic Wind Power Forecasting. Coming soon to  arXiv.
\item  M\o ller, J. K., Zugno, M., \& Madsen, H. (2016). Probabilistic Forecasts of Wind Power Generation by Stochastic Differential Equation Models. Journal of Forecasting, 35(3), 189-205. doi:10.1002/for.2367 
\item  Elkantassi, S., Kalligiannaki, E., \& Tempone, R. (2017). Inference And Sensitivity In Stochastic Wind Power Forecast Models. Proceedings of the 2nd International Conference on Uncertainty Quantification in Computational Sciences and Engineering (UNCECOMP 2017). doi:10.7712/120217.5377.16899
\item Giebel, G., Brownsword, R., Kariniotakis, G., Denhard, M., \& Draxl, C. (2011). The State-Of-The-Art in Short- Term Prediction of Wind Power: A Literature Overview, 2nd edition. ANEMOS.plus. https://doi.org/10.11581/DTU:00000017 
\item Chang, W.-Y. (2014) A Literature Review of Wind Forecasting Methods. Journal of Power and Energy Engineering, 2, 161-168. http://dx.doi.org/10.4236/jpee.2014.24023
\end{itemize}
\end{flushleft}
